### 2020-06-24 TCS_version/08-conclusion.tex
altri piccoli cambiamenti

% !TEX root = ./main_TCS.tex

\section{Conclusion}\label{sec:conclusion}

In this paper we have introduced c\CNA, a variant of the {\tt link}-calculus
where prefixes are link chains and no more single links, as it was 
briefly described in the future work section in~\cite{BBB17}.
This variant allowed us to define 
%shown 
a faithful embedding of 
RSs,
%Reaction Systems,
an emerging formalism to model computationally biochemical systems.
%into multilink processes.  
This translation shows several benefits.
For instance, in our paper the context of a RS can work non deterministically and it 
is recursively defined.
%  the context behaviour is expressed recursively and no be expressed recursively. 
Also, entity mutations can be expressed easily and different reaction systems can
communicate between them.

We have then modified the classical notion (in concurrency) of bisimulation
making it more suitable for proving properties of 
RSs
%Reaction Systems 
in our
framework.
We have defined a new assertion language, which allows us to specify
the properties to be verified, and have shown that it extends Hennessy-Milner logic
to our framework.
We believe that our work can make 
possible to investigate how to integrate our methodology
with other formal techniques to prove 
properties of the modeled systems \cite{CFHOT15,OCHF16,BBGLBH2017}.

We are confident that our embedding can contribute to extend the applications
of 
RSs
%Reaction Systems
to diverse fields of computer science, and life
sciences.

As future work, we plan to implement a prototype of our framework,
with an automatic translation from RSs to  {\tt link}-calculus. 
We can also exploit the implementation of  the symbolic semantics of 
the {\tt link}-calculus~\cite{BrodoO17} that can be found in~\cite{tool}.

We also plan to build on the work in this paper to extend the framework
of RSs towards a model where different RSs are allowed to communicate.