### 2020-07-27 Catuscia - final/04-CCNA.tex

% !TEX root = ./main_RS2L_LNCS.tex

\section{Chained \CNA \ (c\CNA)}
\label{sec:ccna}

In this section we introduce  the syntax and operational semantics
of a variant of the {\tt link}-calculus~\cite{BodeiBB12}, the c\CNA\  (chained \CNA),
where the prefixes are link chains.
%In this section we just focus on the basic ideas of link
% interaction, while we shall enhance the model with name mobility in
% the next section.

\paragraph{Link Chains.}
Let $\mathcal{C}$ be the set of channels, ranged over by $a,b,...$, and 
let $\mathcal{N} = \mathcal{C} \cup \setof{\silent} \cup \setof{{\noacts}}$ be the set of actions, 
ranged over by $\alpha,\beta,...$,
where the symbol $\silent$ denotes a \emph{silent} action, while the symbol $\noact$ denotes a \emph{virtual} (non-specified) action.
A \emph{link} is a pair $\ell=\link{\alpha}{\beta}$;
it is \emph{solid} if $\alpha,\beta \neq \ \noacts$; 
the link $\link{\noact}{\noact}$ is called \emph{virtual};
%\begin{color}{red}
the link $\link{\silent}{\silent}$ is called \emph{silent}.
%\end{color}
A link is \emph{valid} if it is solid or virtual.
%We let $\mathcal{L}$ be the set of valid links.
%
A \emph{link chain} is a 
finite sequence $v = \ell_{1}...\ell_{n}$ of valid links  $\ell_{i} = \link{\alpha_{i}}{\beta_{i}}$ such that:
\begin{enumerate}
\item for any $i\in [1,n-1]$, 
$\left\{\begin{array}{ll}
\beta_{i},\alpha_{i+1}\in \mathcal{C} & \mbox{ implies } \beta_{i} = \alpha_{i+1}\\
\beta_{i}=\silent & \mbox{ iff } \alpha_{i+1}=\silent
\end{array}\right.
$
\item  $\exists i \in [1,n]. \ \ell_i \neq \link{\noact}{\noact}$.
\end{enumerate}

%\begin{color}{red}
A link chain whose links are silent is also called silent.
%\end{color}
Virtual links $\link{\noact}{\noact}$ represent missing elements of a chain. 
The equivalence $\blackstretcheq$ models expansion and contraction of virtual links to adjust the length of a link chain.

\begin{definition}[Equivalence $\blackstretcheq$]\label{def:black}
We let $\blackstretcheq$ be the least equivalence relation 
over link chains closed under the axioms (whenever both sides are well defined):
\[
\begin{array}{rclcrcl}
v\link{\noact}{\noact} & \blackstretcheq &  v & \qquad &
v_1 \startchain{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}v_2 & \blackstretcheq & v_1 \link{\noact}{\noact} v_2\\
\link{\noact}{\noact}v & \blackstretcheq & v & &
v_1 \startchain{\alpha}\chainedlink{a}{a}\chainend{\beta} v_2
& \blackstretcheq & 
v_1 \startchain{\alpha}\chainedlink{a}{\noact}\chainedlink{\noact}{a}\chainend{\beta}v_2 
\end{array}
\]
\end{definition}

Two link chains of equal length can be merged whenever each position occupied by a solid link in one chain is occupied by a virtual link in the other chain and solid links in adjacent positions match. Positions occupied by virtual links in both chains remain virtual. Merging is denoted by $\merges{v_1}{v_2}$.
%
For example, given $v_1 = \startchain{a}\chainedlink{b}{\noact}\chainedlink{\noact}{\noact}\chainend{\noact}$ and $v_2 = \startchain{\noact}\chainedlink{\noact}{b}\chainedlink{c}{\noact}\chainend{\noact}$ we have $\merges{v_1}{v_2} = \startchain{a}\chainedlink{b}{b}\chainedlink{c}{\noact}\chainend{\noact}$.

Some names in a link chain can be restricted as non observable and transformed into silent actions $\silent$. This is possible only if they are matched by some adjacent link. Restriction is denoted by $\restrict{a}v$.
%
For example, given $v = \startchain{a}\chainedlink{b}{b}\chainedlink{c}{\noact}\chainend{\noact}$ we have 
$\restrict{b}v = \startchain{a}\chainedlink{\silent}{\silent}\chainedlink{c}{\noact}\chainend{\noact}$.

\paragraph{Syntax.}
%\begin{definition}
The c\CNA\ processes are generated by the following grammar:
\[
\begin{array}{rcl}
P,Q & ::= &
%\nil \prodsep
%X \prodsep
%s.P  \prodsep
\sum_{i\in I} \upsilon_i.P_i \prodsep
P|Q \prodsep
\restrict{a} P \prodsep
P[\phi] \prodsep
A
\end{array}
\]
\noindent
where $\upsilon_i$ is a  link chain,
%(i.e.\  $\upsilon_i =
%\startchain{\alpha}\chainedlink{\beta}{\noacts}\chainedlink{\noacts}{\alpha'}\chainend{\beta'}$
%with $\alpha,\alpha',\beta,\beta'\neq \ \noacts$),
$\phi$ is a channel renaming function,
and $A$ is a process identifier for which we assume a definition $A
\defeq P$ is available in a given set $\Delta$ of (possibly
recursive) process definitions. We let $\nil$, the inactive process,
denote the empty summation.
%\end{definition}

%\begin{color}{red}
The syntax of c\CNA\  extends that of \CNA~\cite{BBB17} by allowing to use link chains as prefixes instead of links.
This extension was already discussed in~\cite{BBB17} and it preserves all the main formal properties of \CNA. 
For the rest it features nondeterministic choice, parallel composition, restriction, relabelling and possibly recursive definitions of the form $A\defeq P$ for some constant $A$. Here we do not consider name mobility, which is present instead in the {\tt link}-calculus.
%\end{color}

\paragraph{Semantics.}

%\begin{color}{red}
The operational semantics of c\CNA\ is defined in the SOS style by the inference rules in Fig.\ref{fig:cnasos}. The rules are reminiscent of those for Milner's CCS and they essentially coincide with those of \CNA\ in~\cite{BBB17}, except for the presence of prefixes that are link chains instead of single links.
Briefly: rule (\textit{Sum}) selects one alternative and uses, as a label, a possible contraction/expansion $v$ of the link chain $v_j$ in the selected prefix; rule (\textit{Ide}) selects one transition of the process defined by a constant; rule (\textit{Rel}) renames the channels in the label as indicated by $\phi$; rule (\textit{Res}) restricts some names in the label (it cannot be applied when $\restrict{a}v$ is not defined); rules (\textit{Lpar}) and (\textit{Rpar}) account for interleaving in parallel composition; rule (\textit{Com}) 
synchronises interactions (it cannot be applied when $\merges{v}{v'}$ is not defined).
%
Analogously to \CNA, the operational semantics of c\CNA\ satisfies the so called Accordion Lemma: whenever $P \xrightarrow{v} Q$ and $v'\blackstretcheq v$ then $P \xrightarrow{v'} Q$.
%
As a matter of notation, we write $P \rightarrow Q$ when $P \xrightarrow{v} Q$ for some silent link chain $v$ and call it a \emph{silent transition}. Similarly, a sequence of $j$ silent transitions is denoted $P \rightarrow^j Q$.
%\end{color}

%\begin{figure}[t]
%\begin{center} 
%\begin{prooftree} 
%\AxiomC{$\upsilon' \blackstretcheq \upsilon $} %\mbox{($\ell$ {\em only} solid link in $s$)}
%\RightLabel{\scriptsize(\textit{Act})}
%\UnaryInfC{$\upsilon.P \xrightarrow{\upsilon'} P$} 
%% \DisplayProof
%\end{prooftree}
%\end{center}
%\caption{SOS semantic rules of the c\CNA. Only rule $(Act)$ is
%changed. The others are the same as in Figure~\ref{fig:cnasos}, in the Appendix.}
%\label{fig:cnasos}
%\end{figure}

\begin{figure}[t]
\begin{center} 
\begin{prooftree} 
\AxiomC{$v \blackstretcheq v_j\quad j\in I$} %\mbox{($\ell$ {\em only} solid link in $s$)}
\RightLabel{\scriptsize(\textit{Sum})}
\UnaryInfC{$\sum_{i\in I} \upsilon_i.P_i \xrightarrow{v} P_j$} 
\DisplayProof
\
\
\AxiomC{$P \xrightarrow{v} P'$}
\AxiomC{$(A \defeq P)\in\Delta$}
\RightLabel{\scriptsize(\textit{Ide})}
\BinaryInfC{$A \xrightarrow{v} P'$} 
\end{prooftree} 
\end{center}

\begin{center} 
\begin{prooftree} 
\AxiomC{$P \xrightarrow{v} P'$} 
\RightLabel{\scriptsize(\textit{Rel})}
\UnaryInfC{$P[\phi] \xrightarrow{\phi(v)} P'[\phi]$} 
\DisplayProof
\
\
\AxiomC{$P \xrightarrow{v} P'$} 
\RightLabel{\scriptsize(\textit{Res})}
\UnaryInfC{$\restrict{a}P \xrightarrow{\restrict{a}v} \restrict{a}P'$} 
%\AxiomC{$P \xrightarrow{v} P'$} 
%\RightLabel{\scriptsize(\textit{Lsum})}
%\UnaryInfC{$P+Q \xrightarrow{v} P'$} 
%\DisplayProof
%\
%\AxiomC{$Q \xrightarrow{s} Q'$} 
%\RightLabel{\scriptsize(Rsum)}
%\UnaryInfC{$P+Q \xrightarrow{s} Q'$} 
\end{prooftree} 
\end{center}

\begin{center} 
\begin{prooftree} 
%\,
%\AxiomC{$P \xrightarrow{v} P'$} 
%\RightLabel{\scriptsize(\textit{Rel})}
%\UnaryInfC{$P[\phi] \xrightarrow{v[\phi]} P'[\phi]$} 
%\DisplayProof
\AxiomC{$P \xrightarrow{v} P'$} 
\RightLabel{\scriptsize(\textit{Lpar})}
\UnaryInfC{$P|Q \xrightarrow{v} P'|Q$} 
\DisplayProof
\
\
%\AxiomC{$Q \xrightarrow{s} Q'$} 
%\RightLabel{\scriptsize(Rpar)}
%\UnaryInfC{$P|Q \xrightarrow{s} P|Q'$} 
%\end{prooftree} 
%\end{center}
%\begin{center} 
%\begin{prooftree} 
\AxiomC{$P \xrightarrow{v'} P'$} 
\AxiomC{$Q \xrightarrow{v} Q'$} 
\RightLabel{\scriptsize(\textit{Com})}
\BinaryInfC{$P|Q \xrightarrow{\merges{v}{v'}} P'|Q'$} 
%\DisplayProof
\end{prooftree} 
\end{center}
\caption{SOS semantics of the c\CNA\  (rules (\textit{Rel}) and (\textit{Rpar}) omitted).}
\label{fig:cnasos}
\end{figure}

\subsection{Notation for link chains}

Hereafter we make use of some new notations for link chains that will facilitate  
the presentation of our translation.
%First, we define the juxtaposition of link or link chains as
\begin{definition}[Replication]
Let $\upsilon$ be a link chain. Its $n$ times replication $\upsilon^n$ is defined recursively by letting
$\upsilon^0 = \epsilon$ (i.e. the empty chain) and $\upsilon^n = \upsilon^{n-1}\upsilon$,
%\[
%\begin{array}{c@{\qquad}c}
%%\ell ^1 = \ell & \ell^n = \ell^{n-1}\ell\\
%\upsilon^1 = \upsilon& \upsilon^n = \upsilon^{n-1}\upsilon
%\end{array}
%\]
with the hypothesis that all the links in the resulting link chains match.
%, and that $\upsilon^0 = \epsilon$ is the empty chain.
\end{definition}

For example,  the expression $(\startchain{a}\chainedlink{b}{\noact}\chainend{\noact})^3$  denotes the chain $\startchain{a}\chainedlink{b}{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainedlink{\noact}{a}\chainedlink{b}{\noact}\chainend{\noact}$.

 \noindent
We introduce the \emph{half link} that will be used in conjunction with the \emph{open block of chain} to form regular link chains.
\noindent
Let  $\link{a}{}$ denote the \emph{half left link} of a  link $\link{a}{x}$, and conversely
let  $\link{}{a}$ denote the \emph{half right link} of  $\link{x}{a}$.

%\begin{definition}[Half links]
%Let $a$ be a channel name, we define the \emph{half left link}:  $\link{a}{}$,   and the  \emph{half right link}: $\link{}{a}$.
%\end{definition}

%\begin{color}{red}
\begin{definition}[Open block]\label{def:open}
Let $R$ be a totally ordered, finite set of names. We 
define an \emph{open block} as $\blockchain{c_i}{c_o}{c \in R}$\ , where $c_i$ and $c_o$ are annotated version of the name $c$, as follows

\[
\begin{array}{l@{\quad}c@{\quad}l}
\mbox{set}& \mbox{open block expression}& \mbox{result}\\
R= \emptyset & \blockchain{c_i}{c_o}{c \in R}   &\epsilon\\
R = \{ a\}    &\blockchain{c_i}{c_o}{c \in R}  & {}_{a_i}^{\noact}\chainedlink{\noact}{a_o}\\
R = \{a\} \uplus R' \mbox{ with } a=\min R& \blockchain{c_i}{c_o}{c \in R} & {}_{a_i}^{\noact}\chainedlink{\noact}{a_o}\link{}{}\blockchain{c_i}{c_o}{c \in R'}
\end{array}
\]
\end{definition}
%\end{color}
%Please note that, channel names of juxtaposed open blocks of chain are not required to match.

%\noindent
We then combine half links and  open blocks to form regular link chains.
%
%  $\halfl{a}\blockchain{c_i}{c_o}{c \in R}\halfr{b}$ to form a regular link chain; or we can put together half links and open blocks of chain:  $\halfl{a}\blockchain{c_i}{c_o}{c \in R}\blockchain{c_i}{c_o}{c \in R'}\dots \blockchain{c_i}{c_o}{c \in R^{''}}\halfr{b}$ to still form a regular  link chain.
% Again,  all the channel names appearing in the half links or in the open blocks of chain are not required to match.
 For example, for $R = \{a,b\}$ the expression $ \blockchain{c_i}{c_o}{c \in R} $\;
denotes the block of chains ${}_{a_i}^{\noact}\chainedlink{\noact}{a_o} \backslash{}_{b_i}^{\noact}\chainedlink{\noact}{b_o}$; and the expression $\link{r_1}{} \blockchain{c_i}{c_o}{c \in R}\;\link{}{r_2} $ denotes the chain $ \startchain{r_1}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{b_i}{\noact}\chainedlink{\noact}{b_o}\chainend{r_2}$.

%\noindent
%{\color{red} In Definition~\ref{def:open}, we do not care about a specific order for the set $R$, thus we assume to have fixed a specific one.}