### 2020-07-27 Catuscia - final/06-exampleLactose.tex

% !TEX root = ./main_RS2L_LNCS.tex
\section{Example: \emph{lac} operon}
\label{ex:lactose}
In this section we present the encoding of a RS example
taken from~\cite{CMMBM12}.
%, concerning the lac Operon.
\subsection{The \emph{lac} operon}
An operon is a cluster of genes under the control of a single promoter. 
The \emph{lac} operon  is involved in the metabolism of lactose in \emph{Escherichia coli} cells;
it is composed by three adjacent structural genes (plus some regulatory components):  $lacZ$, $lacY$ and $lacA$ encoding for two enzymes $Z$ and $A$, and a transporter $Y$, involved in the digestion of the $lactose$. The main regulations are:
\begin{itemize}
\item 
%\begin{color}{red}
the gene $lacI$ encodes a repressor protein $I$;
%\end{color}
\item the DNA sequence, called \emph{promoter},  is recognised by a RNA polymerase
to iniziate the transcription  of the genes $lacZ$, $lacY$ and $lacA$;
\item a DNA segment, called  the \emph{operator} ($OP$),   obstructs the RNA polymerase functionality when  the repressor protein $I$ is bound to it forming $I\textrm{-}OP$;
\item  a short DNA sequence, called the $CAP\textrm{-}binding\ site$, when it is bound to the complex composed by  the protein $CAP$ and the signal molecule $cAMP$, acts as a promoter for the interaction between the RNA polymerase and the promoter.
\end{itemize}
The functionality of the \emph{lac} operon depends on the integration of two control mechanisms, one mediated by \emph{lactose}, and the other one mediated by \emph{glucose}.

In the first control mechanism, an effect of the absence of the \emph{lactose} is that $I$ is able to bind the operator sequence preventing the \emph{lac} operon expression. If lactose is available, $I$ is unable to bind the operator sequence, and the \emph{lac} operon can be potentially expressed.

In  the second control mechanism, when glucose is absent, the molecule $cAMP$ and the protein $CAP$ increase the \emph{lac} operon expression, thanks to the fact that  the binding between the molecular complex $cAMP\textrm{-}CAP$ and the $CAP\textrm{-}binding\ site$ increases.
In summary, the condition promoting the operon
gene expression is when the $lactose$ is present and the $glucose$ is absent.

In the following we report the description of the \emph{lac} operon mechanism in the reaction system formalism
and then show its encoding in c\CNA.

\subsection{The RS formalization}\label{subsec:RSlacoperon}
The reaction system for the \emph{lac} operon is defined as
$A_{lac} = (S, A)$, where the set S represents
the main biochemical components involved in this genetic system, while the reaction set A contains the biochemical
reactions involved in the regulation of the \emph{lac} operon expression. Formally, the \emph{lac} operon reaction system is defined as
follows:
$S$ is the set
{\small
$$\{lac, Z, Y, A, lacI, I, I\textrm{-}OP, cya, cAMP, crp, CAP,cAMP\textrm{-}CAP, lactose,glucose\},$$
}
and A consists of the following 10 reactions:
{\small
\[
\begin{array}{lcllcl}
a_1 &= &(\{lac\},\{. . .\},\{lac\}), & a_6 &= &(\{cya\},\{. . .\},\{cAMP\}),\\
a_2 &= &(\{lacI\},\{. . .\},\{lacI\}), & a_7 &= &(\{crp\},\{. . .\},\{crp\}),\\
a_3 &= &(\{lacI\},\{. . .\},\{I\}), & a_8 &= &(\{crp\},\{. . .\},\{CAP\}),\\
a_4 &= &(\{I\},\{lactose\},\{I\textrm{-}OP\}), & a_9 &= &(\{cAMP, CAP\},\{glucose\},\{cAMP\textrm{-}CAP\}),\\
a_5 &= &(\{cya\},\{. . .\},\{cya\}),&
a_{10} &=& (\{lac, cAMP\textrm{-}CAP\},\{I\textrm{-}OP\},\{Z, Y, A\}).
\end{array}
\]
}
The default context ($DC$) is composed by those entities that are always present in the system
$DC = \{lac,lacI,I,cya,cAMP,crp,CAP\}$, whereas the $lactose$ and the $glucose$ are given non-deterministically by the context.
\subsection{The RS encoding}
For the sake of readability, the encoding we propose exploits the specific features of the example in hand 
to perform some simplifications:
\begin{itemize}
\item for the entities in the default context, $s \in DC$, as they are persistent, we do not provide the $\overline{P_s}$ processes and the $Cxt_s$ processes;
\item for the reactions requiring the presence of entities $s \in DC$, we do not provide the reaction alternative behaviour when $s$ is absent;
\item the $Cxt_s$ processes are specified only for those entitiess that are really provided by the context.
\end{itemize}
Moreover, we do not model the \emph{dummy} entity that is specified by  dots ($\dots$) by the RS reactions in Section~\ref{subsec:RSlacoperon}.
Finally, we exclude the \emph{duplication reactions} ($a_1$, $a_2$, $a_5$, $a_7$), and renumber the remaining reactions :
{\small
\[
\begin{array}{cccl}
%a1 &= &(\{lac\},\{. . .\},\{lac\}),\\
%a2 &= &(\{lacI\},\{. . .\},\{lacI\}),\\
\textrm{old } &\textrm{new }&& \textrm{reactions}\\
a_3 &a_1&=&(\{lacI\},\{. . .\},\{I\}),\\
a_4 &a_2&=&(\{I\},\{lactose\},\{I\textrm{-}OP\}),\\
%a5 &= &(\{cya\},\{. . .\},\{cya\}),\\
a_6 & a_3&=&(\{cya\},\{. . .\},\{cAMP\}),\\
%a7 &= &(\{crp\},\{. . .\},\{crp\}),\\
a_8 &a_4&=&(\{crp\},\{. . .\},\{CAP\}),\\
a_9 &a_5&=&(\{cAMP, CAP\},\{glucose\},\{cAMP\textrm{-}CAP\}),\\
a_{10} & a_6&=& (\{lac, cAMP\textrm{-}CAP\},\{I\textrm{-}OP\},\{Z, Y, A\}).
\end{array}
\]
}
%\paragraph{Duplication reactions.}
%As explained previously, their encoding is omitted.
%%$
%\begin{array}{lcccl}
%P_{ai} & \defeq& P_d(ai,s) & = & \startchain{r_i}\chainedlink{s_i}{\noact}\chainedlink{\noact}{s_o}
%\chainedlink{\tilde{s}_i}{\noact}\chainedlink{\noact}{\tilde{s}_o}\chainend{r_{i+1}}.P_d(ai,s)
%\end{array}
%$,
%with $r_1 = \silent$.
%There are four duplication rules:\\ $P_d(a1,lac)$, $P_d(a2,lacI)$,$P_d(a5,cya)$,$P_d(a7,crp)$.\\
\paragraph{Expression reactions.}
First we define the parametric process
\[
\begin{array}{lcl}
% P_i(s1,s2) & \defeq & \startchain{r_i}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}
%\chainedlink{\tilde{s2}_i}{\noact}\chainedlink{\noact}{\tilde{s2}_o}\chainend{r_{i+1}}.P_i(s1,s2)\\
 P_i(s1,s2) & \defeq & \startchain{r_i}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}
 \chainedlink{r_{i+1}}{\noact}\chainedlink{\noact}{p_{i}}\chainedlink{\tilde{s2}_i}{\noact}
\chainedlink{\noact}{\tilde{s2}_o}\chainend{p_{i+1}}.P_i(s1,s2)\\
\end{array}
\]
Then, we let $P_{a1}\defeq P_1(lacI,I)$,  $P_{a3} \defeq P_3(cya,cAMP)$, and $P_{a4} \defeq P_4(crp,CAP)$.
%\[
%\begin{array}{lcl}
%P_{a1} & =& \startchain{\silent}\chainedlink{lac_i}{\noact}\chainedlink{\noact}{lac_o}\chainedlink{\tilde{lac}_i}{\noact}\chainedlink{\noact}{\tilde{lac}_o}\chainend{r_2}.P_{a1}\\
%P_{a2} & =& \startchain{r_2}\chainedlink{lacI_i}{\noact}\chainedlink{\noact}{lacI_o}\chainedlink{\tilde{lacI}_i}{\noact}\chainedlink{\noact}{\tilde{lacI}_o}\chainend{r_3}.P_{a2}\\
%P_{a3} & =& \startchain{r_3}\chainedlink{I_i}{\noact}\chainedlink{\noact}{I_o}\chainedlink{\tilde{I}_i}{\noact}\chainedlink{\noact}{\tilde{I}_o}\chainend{r_4}.P_{a3}\\
\paragraph{Regulation reactions.}

{\small
\[
\begin{array}{lcl}
%P_{a2} & \defeq& \startchain{r_2}\chainedlink{I_i}{\noact}\chainedlink{\noact}{I_o}\chainedlink{\overline{lactose}_i}{\noact}\chainedlink{\noact}{\overline{lactose}_o}
%\chainedlink{\widetilde{I\textrm{-}OP}_i}{\noact}\chainedlink{\noact}{\widetilde{I\textrm{-}OP}_o}\chainend{r_3}.P_{a2}  \\
P_{a2} & \defeq& \startchain{r_2}\chainedlink{I_i}{\noact}\chainedlink{\noact}{I_o}\chainedlink{\overline{lactose}_i}{\noact}\chainedlink{\noact}{\overline{lactose}_o}
\chainedlink{r_3}{\noact}\chainedlink{\noact}{p_2}
\chainedlink{\widetilde{I\textrm{-}OP}_i}{\noact}\chainedlink{\noact}{\widetilde{I\textrm{-}OP}_o}\chainend{p_3}.P_{a2}  \\
&&+\\
 && \startchain{r_2}\chainedlink{lactose_i}{\noact}\chainedlink{\noact}{lactose_o}\chainedlink{r_3}{\noact}\chainedlink{\noact}{p_2}\chainend{p_3}.P_{a2} \ +\  \startchain{r_2}\chainedlink{\overline{I}_i}{\noact}\chainedlink{\noact}{\overline{I}_o}\chainedlink{r_3}{\noact}\chainedlink{\noact}{p_2}\chainend{p_3}.P_{a2}\\[5pt]
 \\
%%P_{a5} & =& \startchain{r_5}\chainedlink{cya_i}{\noact}\chainedlink{\noact}{cya_o}\chainedlink{\tilde{cya}_i}{\noact}\chainedlink{\noact}{\tilde{cya}_o}\chainend{r_6}.P_{a5}\\
%P_{a6} & =& \startchain{r_6}\chainedlink{cya_i}{\noact}\chainedlink{\noact}{cya_o}\chainedlink{\tilde{cya}_i}{\noact}\chainedlink{\noact}{\tilde{cya}_o}\chainend{r_7}.P_{a6}\\
%%P_{a7} & =& \startchain{r_7}\chainedlink{crp_i}{\noact}\chainedlink{\noact}{crp_o}\chainedlink{\tilde{crp}_i}{\noact}\chainedlink{\noact}{\tilde{crp}_o}\chainend{r_8}.P_{a7}\\
%P_{a8} & =& \startchain{r_8}\chainedlink{crp_i}{\noact}\chainedlink{\noact}{crp_o}\chainedlink{\tilde{CAP}_i}{\noact}\chainedlink{\noact}{\tilde{CAP}_o}\chainend{r_9}.P_{a8}\\
P_{a5} & \defeq& \startchain{r_5}\chainedlink{cAMP_i}{\noact}\chainedlink{\noact}{cAMP_o}
\chainedlink{CAP_i}{\noact}\chainedlink{\noact}{CAP_o}\chainedlink{\overline{glucose}_i}{\noact}
\chainedlink{\noact}{\overline{glucose}_o}\chainedlink{r_6}{\noact}
\chainedlink{\noact}{p_5}
\chainedlink{\widetilde{cAMP\textrm{-}CAP}_i}{\noact}\chainedlink{\noact}{\widetilde{cAMP\textrm{-}CAP}_o}
\chainend{p_6}.P_{a5}\\
%P_{a5} & \defeq& \startchain{r_5}\chainedlink{cAMP_i}{\noact}\chainedlink{\noact}{cAMP_o}
%\chainedlink{CAP_i}{\noact}\chainedlink{\noact}{CAP_o}\chainedlink{\overline{glucose}_i}{\noact}
%\chainedlink{\noact}{\overline{glucose}_o}
%\chainedlink{\widetilde{cAMP\textrm{-}CAP}_i}{\noact}\chainedlink{\noact}{\widetilde{cAMP\textrm{-}CAP}_o}
%\chainend{r_6}.P_{a5}\\
&& +\\
&& \startchain{r_5}\chainedlink{glucose_i}{\noact}\chainedlink{\noact}{glucose_o} \chainedlink{r_6}{\noact}\chainedlink{\noact}{p_5}\chainend{p_6}.P_{a5}\\
&& + \\
&&  \startchain{r_5}\chainedlink{\overline{cAMP}_i}{\noact}\chainedlink{\noact}{\overline{cAMP}_o} \chainedlink{r_6}{\noact}\chainedlink{\noact}{p_5}\chainend{p_6}.P_{a5} \ + \ \startchain{r_5}\chainedlink{\overline{CAP}_i}{\noact}\chainedlink{\noact}{\overline{CAP}_o} \chainedlink{r_6}{\noact}\chainedlink{\noact}{p_5}\chainend{p_6}.P_{a5}\\[5pt]
\\
%\end{array}
%\]\\
%%$lac$ operon expression.\\
%$
%\begin{array}{lcl}
P_{a6} & \defeq& \startchain{r_{6}}\chainedlink{lac_i}{\noact}\chainedlink{\noact}{lac_o}
\chainedlink{cAMP\textrm{-}CAP_i}{\noact}\chainedlink{\noact}{cAMP\textrm{-}CAP_o}\chainedlink{\overline{I\textrm{-}OP}_i}{\noact}
\chainedlink{\noact}{\overline{I\textrm{-}OP}_o}\chainedlink{cxt_1}{\noact}
\chainedlink{\noact}{p_6}
\chainedlink{\tilde{z}_i}{\noact}\chainedlink{\noact}{\tilde{z}_o}
\chainedlink{\tilde{y}_i}{\noact}\chainedlink{\noact}{\tilde{y}_o}
\chainedlink{\tilde{A}_i}{\noact}\chainedlink{\noact}{\tilde{A}_o}
\chainend{\silent}.P_{a6}\\
%P_{a6} & \defeq& \startchain{r_{6}}\chainedlink{lac_i}{\noact}\chainedlink{\noact}{lac_o}
%\chainedlink{cAMP_i}{\noact}\chainedlink{\noact}{cAMP_o}\chainedlink{\overline{I\textrm{-}OP}_i}{\noact}
%\chainedlink{\noact}{\overline{I\textrm{-}OP}_o}
%\chainedlink{\tilde{z}_i}{\noact}\chainedlink{\noact}{\tilde{z}_o}
%\chainedlink{\tilde{y}_i}{\noact}\chainedlink{\noact}{\tilde{y}_o}
%\chainedlink{\tilde{A}_i}{\noact}\chainedlink{\noact}{\tilde{A}_o}
%\chainend{cxt_1}.P_{a6}\\
&& +\\
&& \startchain{r_{6}}\chainedlink{I\textrm{-}OP_i}{\noact}\chainedlink{\noact}{I\textrm{-}OP_o} \chainedlink{cxt_1}{\noact}\chainedlink{\noact}{p_6}\chainend{\silent}.P_{a6} \\
&& + \\
&&  \startchain{r_{6}}\chainedlink{\overline{lac}_i}{\noact}\chainedlink{\noact}{\overline{lac}_o} \chainedlink{cxt_1}{\noact}\chainedlink{\noact}{p_6}\chainend{\silent}.P_{a6} \ + \ \chainedlink{\overline{cAMP\textrm{-}CAP}_i}{\noact}\chainedlink{\noact}{\overline{cAMP\textrm{-}CAP}_o} \chainedlink{cxt_1}{\noact}\chainedlink{\noact}{p_6}\chainend{\silent}.P_{a6}
\end{array}
\]
}
\paragraph{Processes for the entities.}
We exploit the specificity of the example in hand to optimise the code, and we specify exactly
the number of solid links that each process encoding an entity must offer.
For the always present entities  we let:
%{\color{red} For the always present entities we define the parametric process $P(s)\defeq \link{s_i}{s_o}.P_{s}$, and we let:
%\[
%\begin{array}{lcl@{\qquad}lcl@{\qquad}lcl@{\qquad}lcl}
%P_{cya}&\defeq&P(cya) &P_{crp}&\defeq&P(crp)&P_{lacI}&\defeq&P(lacI) &
% P_{lac}&\defeq&P(lac)
%\end{array}\]
\[
\begin{array}{lcl@{\qquad}lcl}
P_{cya}&\defeq&\link{cya_i}{cya_o}.P_{cya} &
P_{crp}&\defeq&\link{crp_i}{crp_o}.P_{crp}\\[5pt]
P_{lacI}&\defeq& \link{lacI_i}{lacI_o}.P_{lacI} &
 P_{lac}&\defeq&\link{lac_i}{lac_o}.P_{lac}
\end{array}\]
For the entities always produced (i.e. not present only at the first step), we  provide a  parametric
definition  
$P_e(s) \defeq  \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{\tilde{s}_i}\chainend{\tilde{s}_o}.P_e(s) +  \startchain{\tilde{s_i}}\chainend{\tilde{s}_o} .P_e(s) $.\\
% There are four entities of the first type: $P_d(lac)$, $P_d(cya)$, $P_d(crp)$, $P_d(lacI)$.
 There are three entities of the second type:  $$P_{cAMP}\defeq P_e(cAMP) \qquad P_{CAP} \defeq P_e(CAP)\qquad P_I \defeq P_e(I).$$

The entity $I\textrm{-}OP$ can be either produced (by $a_2$) or tested for absence (by $a_6$).
Correspondingly, the process  $P_{I\textrm{-}OP}$ is defined as follows:
\[\begin{array}{lcl}
P_{I\textrm{-}OP}&\defeq & 
\sum_{h=0}^1 (\startchain{I\textrm{-}OP_i}\chainedlink{I\textrm{-}OP_o}{\noact}\chainend{\noact})^h  \, \startchain{\widetilde{I\textrm{-}OP}_i}\chainend{\widetilde{I\textrm{-}OP}_o}.P_{I\textrm{-}OP}
%\\
%&&+\\
%&&\sum_{h=0}^1(\startchain{I-OP_i}\chainedlink{I-OP_o}{\noact}\chainend{\noact})^h\ \startchain{\widetilde{I-OP}_i}\chainedlink{\widetilde{I-OP}_o}{\noact}\chainedlink{\noact}{\underline{I-OP}_i}\chainend{\underline{I-OP}_o}.P_{I-OP}\\
%&&
\  +\ %\\
%&&
\link{I\textrm{-}OP_i}{I\textrm{-}OP_o}.\overline{P_{I\textrm{-}OP}}\\
\overline{P_{I\textrm{-}OP}} & \defeq & 
% P_{I\textrm{-}OP}(\overline{I\textrm{-}OP})\\
\sum_{h=0}^1  (\startchain{\overline{I\textrm{-}OP}_i}\chainedlink{\overline{I\textrm{-}OP}_o}{\noact}\chainend{\noact})^h \,
\startchain{\widetilde{I\textrm{-}OP}_i}\chainend{\widetilde{I\textrm{-}OP}_o}.P_{I\textrm{-}OP}%\\
%&&
\ +\ %\\
%&&
\link{\overline{I\textrm{-}OP}_i}{\overline{I\textrm{-}OP}_o}.\overline{P_{I\textrm{-}OP}}\\

\end{array}
\]
\noindent
The  process $P_{cAMP\textrm{-}CAP}$ is similar to $P_{I\textrm{-}OP}$, as it is produced by $a_5$ and tested for presence by $a_6$. 
Its code is omitted.
%Its code is  in Table~\ref{tab:cAMP-CAP}, in the Appendix.

The $lactose$ is provided by the context and tested for absence by $a_2$.

\[
\begin{array}{lcllcl}
P_{lactose} &\defeq &%== \overline{P_{glucose}}&\defeq & P_{glu-lact}(\overline{glucose})\\
\sum_{h=0}^1 (\startchain{lactose_i}\chainedlink{lactose_o}{\noact}\chainend{\noact})^h\ \link{\widehat{lactose_i}}{\widehat{lactose_o}}.P_{lactose}
\\
&& + \\
&& \sum_{h=0}^1 (\startchain{lactose_i}\chainedlink{lactose_o}{\noact}\chainend{\noact})^h\ \link{\underline{lactose_i}}{\underline{lactose_o}}.\overline{P_{lactose}}\\
\\
\overline{P_{lactose}} &\defeq &\sum_{h=0}^1 (\startchain{\overline{lactose}_i}\chainedlink{\overline{lactose}_o}{\noact}\chainend{\noact})^h\ \link{\widehat{lactose_i}}{\widehat{lactose_o}}.P_{lactose}\\
&&+\\
&& \sum_{h=0}^1 (\startchain{\overline{lactose}_i}\chainedlink{\overline{lactose}_o}{\noact}\chainend{\noact})^h\ \link{\underline{lactose_i}}{\underline{lactose_o}}.\overline{P_{lactose}}\\
\end{array}
\]
\noindent
The  process $P_{glucose}$ is similar to $P_{lactose}$ and tested for absence by $a_5$. 
Its code is omitted.
%Its code is  in Table~\ref{tab:glucose}, in the Appendix.

The entity $z$ can only be produced by rule $a_6$, while it is never provided by the context. Moreover, there is no rule for testing its presence or absence.
\[
\begin{array}{lcl@{\qquad\qquad}lcl}
P_z&\defeq &\startchain{\tilde{z_i}}\chainedlink{\tilde{z_o}}{\noact}\chainedlink{\noact}{\underline{z_i}}\chainend{\underline{z_o}}.P_z\ + \
\link{\underline{z_i}}{\underline{z_o}}.\overline{P_z}
&
\overline{P_z} &\defeq &\startchain{\tilde{z_i}}\chainedlink{\tilde{z_o}}{\noact}\chainedlink{\noact}{\underline{z_i}}\chainend{\underline{z_o}}.\overline{P_z}\ + \
\link{\underline{z_i}}{\underline{z_o}}.\overline{P_z}\\\end{array}
\]
\noindent
The entities $y$ and $A$ are treated likewise $z$. Their processes are omitted.
%in Table~\ref{tab:yA} in the Appendix.

\paragraph{Context.}
%Now, we define the behaviour of the context.
The entities in $DC$ are assumed always present by default, so no context process is needed for them.
The entities $z$, $y$, and $A$ are assumed never provided by the context:
% Their processes are
\[
\begin{array}{rcl@{\qquad}rcl}
Cxt_{z}&\defeq & \startchain{cxt_1}\chainedlink{\underline{z}_i}{\noact}\chainedlink{\noact}{\underline{z}_o}\chainend{cxt_2}.Cxt_{z}
&
Cxt_{y}&\defeq & \startchain{cxt_2}\chainedlink{\underline{y}_i}{\noact}\chainedlink{\noact}{\underline{y}_o}\chainend{cxt_3}.Cxt_{y}
\\
Cxt_{A}&\defeq & \startchain{cxt_3}\chainedlink{\underline{A}_i}{\noact}\chainedlink{\noact}{\underline{A}_o}\chainend{cxt_4}.Cxt_{A} 
\end{array}
\]
 Also, for the sake of presentation, we assume that the $glucose$ is never provided and $lactose$ is always provided by the context:
\[
\begin{array}{lcl}
Cxt_{lactose}&\defeq& \startchain{cxt_4}\chainedlink{\widehat{lactose}_i}{\noact}\chainedlink{\noact}{\widehat{lactose}_o}\chainend{cxt_5}.Cxt_{lactose}\\
Cxt_{glucose}&\defeq & \startchain{cxt_5}\chainedlink{\underline{glucose}_i}{\noact}\chainedlink{\noact}{\underline{glucose}_o}\chainend{p_1}.Cxt_{glucose}
\end{array}
\]

\noindent
%For the sake of presentation, we assume that the $lactose$ is always provided by the context, in contrast,
%$glucose$ is never provided.
%\[
%\begin{array}{lcllcl}
%Cxt_{lactose}&\defeq& \startchain{cxt_4}\chainedlink{\widehat{lactose}_i}{\noact}\chainedlink{\noact}{\widehat{lactose}_o}\chainend{cxt_5}.Cxt_{lactose}
%\\
%Cxt_{glucose}&\defeq & \startchain{cxt_5}\chainedlink{\underline{glucose}_i}{\noact}\chainedlink{\noact}{\underline{glucose}_o}\chainend{p_1}.Cxt_{glucose}
%\end{array}
%\]
\noindent
In the following we let $CXT \defeq Cxt_z | Cxt_y |Cxt_A | Cxt_{lactose} | Cxt_{glucose}$ be the processes for context. 
%
The whole system
%, encoding an extended interactive process, with the previous context behaviour
is as follows: 
\[ lacOp \defeq \restrict{names} (\Pi_{i=1}^6 P_{ai} | \Pi_{s \in DC} P_s |   \Pi_{s \in S \backslash DC} \overline{P}_s| CXT)
\]

\paragraph{Execution.}
Now, we show  the execution of two transitions. 
After the first transition, the entity $cAMP\textrm{-}CAP$ is produced due to the absence of  $glucose$(see $P_{a5}$), while the presence of $lactose$ inhibits the production of $I\textrm{-}OP$ (see $P_{a2}$): $lacOp \xrightarrow{\restrict{names} v} lacOp'$,
where
{\small
$$
\begin{array}{rcl}
v & = & \link{\silent}{lac_i}\dots \link{lactose_o}{r3}\dots
 \link{r_5}{cAMP_i}\dots
  \link{\overline{glucose}_o}{r_6}
\dots \link{p_5}{\widetilde{cAMP\textrm{-}CAP}_i}\dots \link{p_6}{\silent}\\[5pt]
%{\underline{glucose}_o}
lacOp' & \defeq & \restrict{names} (\Pi_{i =1}^6 P_{ai} | \Pi_{s \in AP} P_s |   \Pi_{s \in S \backslash AP} \overline{P}_s| CXT)
\end{array}
$$ 
}
with $AP= DC\cup \{cAMP\textrm{-}CAP\}$  the actual context.\\
After the second step the entities $z$, $y$ and $A$ are produced, due to the presence of $cAMP\textrm{-}CAP$ and the absence of $I\textrm{-}OP$ (see $P_{a6}$), thus
$lacOp' \xrightarrow{\restrict{names}v'} lacOp''$
where:
{\small
$$
\begin{array}{rcl}
v' & = & \link{\silent}{lac_i}\dots \link{lac_o}{cAMP-CAP_i}\dots \startchain{\overline{I-OP}_o}
\chainedlink{cxt_1}{\dots}\startchain{\underline{glucose}_o}\chainedlink{p_1}{p_1}
\chainedlink{\dots}{p_6}\chainedlink{\tilde{z}_i}{\tilde{z}_i}\chainedlink{\tilde{z}_o}{\tilde{z}_o}
\chainedlink{\tilde{y}_i}{\tilde{y}_i}\chainedlink{\tilde{y}_o}{\tilde{y}_o}\chainedlink{\tilde{A}_i}{\tilde{A}_i}\chainedlink{\tilde{A}_o}{\tilde{A}_o}\chainend{\silent}\\
%\chainend{cxt_1}\dots \link{\underline{glucose}_o}{\silent}\\[5pt]
lacOp'' & \defeq & \restrict{names} (\Pi_{i =1}^{6} P_{ai} | \Pi_{s \in AP'} P_s |   \Pi_{s \in S \backslash AP'} \overline{P}_s| CXT)
\end{array}
$$
}
with $AP'= DC\cup \{z,y,A\} $.