### 2020-07-27 fragments/old/05-translation_OLD.tex

% !TEX root = ./main_RS2L.tex

\section{From Reaction Systems to c\CNA}
\label{sec:trans}

Here we present a translation from the Reaction Systems~\cite{} to a variant of the \CNA, the c\CNA, where
the prefixes are link chains and no simple links.

\begin{definition}
\label{def:trans}
Let ${\cal A} = (S, A)$ be a reaction system, then its  translation, $P=[|{\cal A}|]$ into the {\tt link}-calculus is defined as follows 
\begin{center}$P = \restrict{stop,r_1,\dots,r_m}(\Pi_{s\in S} P_s | \Pi_{i\in A} P_a| STOP)$,\end{center} 
with $m$ be the number of reactions, and where:
\begin{itemize}
\item for each biological element $s \in S$ we write:\\
\[
\begin{array}{lcl}
%P_s& \defeq &\sum_{h=1}^{{\cal N}_s}\underbrace{ \startchain{a_i}\chainedlink{a_o}{\noact}\chainedlink{\noact}{a_i}\chainedlink{a_o}{\noact}\chainend{\noact}...}_{size = h}.\nil + 
%\\
%&&| \ \\
%&& (  \sum_{h=1}^{{\cal P}_s}\underbrace{ \startchain{\tilde{a}_i}\chainedlink{\tilde{a}_o}{\noact}\chainedlink{\noact}{\tilde{a}_i}\chainedlink{\tilde{a}_o}{\noact}\chainend{\noact}...}_{size = h}\link{\hat{a}_i}{\hat{a}_o}.P_s
% \ + \ \link{\hat{a}_i}{\hat{a}_o}.\overline{P_s})
\\
P_s& \defeq &\sum_{h\geq 0} \sum_{k\geq 1}  \underbrace{ \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{h\ \mathrm{times}}\underbrace{ \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainedlink{\noact}{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact}...}_{k\ \mathrm{times}}\link{\hat{s}_i}{\hat{s}_o}.P_s\\
&&\ +\\
&&\sum_{h\geq 0}   \underbrace{ \startchain{s_i}\chainedlink{s_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s_o}{\noact}\chainend{\noact}...}_{h\ \mathrm{times}}\link{\hat{s}_i}{\hat{s}_o}.\overline{P}_s

% \ + \ \link{\hat{a}_i}{\hat{a}_o}.\overline{P_s})
 \\
 \overline{P}_s& \defeq &\sum_{h\geq 0} \sum_{k\geq 1} \underbrace{ \startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact}...}_{h\ \mathrm{times}}\underbrace{ \startchain{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainedlink{\noact}{\tilde{s}_i}\chainedlink{\tilde{s}_o}{\noact}\chainend{\noact}...}_{k\ \mathrm{times}}.\link{\hat{s}_i}{\hat{s}_o}.P_s\\
 && \ +\\
 && \sum_{h\geq 0}  \underbrace{ \startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainend{\noact}...}_{h\ \mathrm{times}}.\link{\hat{s}_i}{\hat{s}_o}.\overline{P}_s\\
 %\ + \ \link{\hat{a}_i}{\hat{a}_o}.\overline{P_s})

\end{array}
\]
\end{itemize}
\begin{itemize}
\item we number each rule in a way each $a \in A$ has associated a progressive number $j$. Then, for each rule $a \in A$ with number $j$, we write:\\
 \[
%\begin{array}{lcl}
%R_i& \defeq & \startchain{rule_i}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{\overline{a}_i}{\noact} \chainedlink{\noact}{\overline{a}_o}\chainedlink{\tilde{a}_i}{\noact}\chainedlink{\noact}{\tilde{a}_o} \chainend{rule_{i+1}}.R_i\\
%&& +\  \\
%&& \startchain{rule_i}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{a_i}{\noact} \chainedlink{\noact}{a_o} \chainend{rule_{i+1}}.R_i\\
%&&+\\
%&& \startchain{rule_i}\chainedlink{\overline{a}_i}{\noact}\chainedlink{\noact}{\overline{a}_o}\chainedlink{\overline{a}_i}{\noact} \chainedlink{\noact}{\overline{a}_o} \chainend{rule_{i+1}}.R_i\\
%&& +\\
%&& \startchain{rule_i}\chainedlink{\overline{a}_i}{\noact}\chainedlink{\noact}{\overline{a}_o}\chainedlink{a_i}{\noact} \chainedlink{\noact}{a_o} \chainend{rule_{i+1}}.R_i\\
%\end{array}
\begin{array}{lcl}
A_j& \defeq & \startchain{r_j}
      \underbrace{\chainedlink{s_{1i}}{\noact}\chainedlink{\noact}{s_{1o}}\chainedlink{s_{2i}}{\noact}\ldots\chainedlink{\noact}{s_{ko}}}_{R_j=\{s_{1},...,s_{k}\}}
      \underbrace{\chainedlink{\overline{b}_{1i}}{\noact} \chainedlink{\noact}{\overline{b}_{1o}}
      \chainedlink{\overline{b}_{2i}}{\noact} \ldots \chainedlink{\noact}{\overline{b}_{ho}}}_{I_j=\{b_1,...,b_h\}}
      \underbrace{\chainedlink{\tilde{c}_{1i}}{\noact}\chainedlink{\noact}{\tilde{c}_{1o}} \chainedlink{\tilde{c}_{2i}}{\noact}\ldots \chainedlink{\noact}{\tilde{c}_{\ell o}}}_{P_j=\{c_1,...,c_\ell\}}
      \chainend{r_{j+1}}.A_j\\
&& +\  \\
&& \sum_{s\in R_j}\startchain{r_j}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o} \chainend{r_{j+1}}.A_j\\
&&+\\
&& \sum_{b\in I_j}\startchain{r_j}\chainedlink{b_i}{\noact} \chainedlink{\noact}{b_o} \chainend{r_{i+1}}.A_i\\
%&& +\\
%&& \startchain{rule_i}\chainedlink{\overline{a}_i}{\noact}\chainedlink{\noact}{\overline{a}_o}\chainedlink{a_i}{\noact} \chainedlink{\noact}{a_o} \chainend{rule_{i+1}}.R_i\\
\end{array}
\]
\end{itemize}
where $s$ is the meta-variable for the elements required by the rule,  $b$ is the meta-variable for  the elements that the rule requires to be absent, and  $c$ is the metavariable for the elements produced.

For notational convenience, we fix that $r_1 = \tau$ and $r_{m+1} = stop$ for $m$ the number of reactions.

Eventually, the unique process $STOP$ is defined as\\
\[
\begin{array}{lcl}
STOP &\defeq& \startchain{stop}\chainedlink{\hat{s1}_{i}}{\noact}\chainedlink{\noact}{\hat{s1}_{o}}\chainedlink{\hat{s2}_{i}}{\noact}... \chainedlink{\noact}{\hat{sn}_{o}} \chainend{\silent}. STOP
\end{array}
\]
where the $n \in \mathbb{N}$ is the number of the elements in the system.
\end{definition}

 \begin{example}
 Let  $\mathcal{A} =(S,A) $ be a reaction system with $S= \{e1,e2,e3\}$, and $A=\{(e1|e3|e1),(e2|e1|e3),(e3|e2|e2) \}$.
 The code of the resulting {\tt link}-calculus process $ \Pi_{s\in S} P_s | \Pi_{i\in A} P_a| STOP = [|(S,A) |] $ is the following. \\
 The code for the elements:\\
%  \[
% \begin{array}{lcl}
% S1 &\defeq& (\startchain{e1_i}\chainedlink{e1_o}{\noact}\chainedlink{\noact} \\
%  +\startchain{e1_i}\chainedlink{e1_o}{\noact}\chainedlink{\noact}{e1_i}\chainend{e1_o}.\nil) \  |\ 
% (\startchain{\tilde{e1}_i}\chainedlink{\tilde{e1}_o}{\noact}\chainedlink{\noact}{\hat{e1}_i}\chainend{\hat{e1}_o}.S1 + \link{\hat{e1}_i}{\hat{e1}_o}.\overline{S1})\\
%\\
%\overline{S1} &\defeq& (\link{\overline{e1}_i}{{e1}_o}.\nil +\startchain{\overline{e1}_i}\chainedlink{\overline{e1}_o}{\noact}\chainedlink{\noact}{\overline{e1}_i}\chainend{\overline{e1}_o}.\nil) \  |\ 
% (\startchain{\tilde{e1}_i}\chainedlink{\tilde{e1}_o}{\noact}\chainedlink{\noact}{\hat{e1}_i}\chainend{\hat{e1}_o}.S1 + \link{\hat{e1}_i}{\hat{e1}_o}.\overline{S1})\\ 
% \\
% \\
%  S2 &\defeq& (\link{e2_i}{e2_o}.\nil +\startchain{e2_i}\chainedlink{e2_o}{\noact}\chainedlink{\noact}{e2_i}\chainend{e2_o}.\nil) \  |\ 
% (\startchain{\tilde{e2}_i}\chainedlink{\tilde{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.S1 + \link{\hat{e2}_i}{\hat{e2}_o}.\overline{S2})\\
%\\
%\overline{S2} &\defeq& (\link{\overline{e2}_i}{{e2}_o}.\nil +\startchain{\overline{e2}_i}\chainedlink{\overline{e2}_o}{\noact}\chainedlink{\noact}{\overline{e2}_i}\chainend{\overline{e2}_o}.\nil) \  |\ 
% (\startchain{\tilde{e2}_i}\chainedlink{\tilde{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.S1 + \link{\hat{e2}_i}{\hat{e2}_o}.\overline{S2})\\ 
% \\
% \\
%   S3 &\defeq& (\link{e3_i}{e3_o}.\nil +\startchain{e3_i}\chainedlink{e3_o}{\noact}\chainedlink{\noact}{e3_i}\chainend{e3_o}.\nil) \  |\ 
% (\startchain{\tilde{e3}_i}\chainedlink{\tilde{e3}_o}{\noact}\chainedlink{\noact}{\hat{e3}_i}\chainend{\hat{e3}_o}.S1 + \link{\hat{e3}_i}{\hat{e3}_o}.\overline{S3})\\
%\\
%\overline{S3} &\defeq& (\link{\overline{e3}_i}{{e3}_o}.\nil +\startchain{\overline{e3}_i}\chainedlink{\overline{e3}_o}{\noact}\chainedlink{\noact}{\overline{e3}_i}\chainend{\overline{e3}_o}.\nil) \  |\ 
% (\startchain{\tilde{e3}_i}\chainedlink{\tilde{e3}_o}{\noact}\chainedlink{\noact}{\hat{e3}_i}\chainend{\hat{e3}_o}.S1 + \link{\hat{e3}_i}{\hat{e3}_o}.\overline{S3})\\ 
% \end{array}
% \]

 \[
 \begin{array}{lcl}
 S1 &\defeq& \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.S1 \ + \
 \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.S1 \ + \ \startchain{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.S1\\
 && + \\
 &&  \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.\overline{S1} 
\  + \ \startchain{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{s1_i}\chainedlink{s1_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.\overline{S1}\ + \  \link{\hat{s1}_i}{\hat{s1}_o}.\overline{S1}\\
\\
\overline{S1} &\defeq& \startchain{\overline{s1}_i}\chainedlink{\overline{s1}_o}{\noact}\chainedlink{\noact}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.S1 \ + \
 \startchain{\overline{s1}_i}\chainedlink{\overline{s1}_o}{\noact}\chainedlink{\noact}{\overline{s1}_i}\chainedlink{\overline{s1}_o}{\noact}\chainedlink{\noact}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.S1 \ + \ \startchain{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.S1\\
 && + \\
 &&  \startchain{\overline{s1}_i}\chainedlink{\overline{s1}_o}{\noact}\chainedlink{\noact}{\overline{s1}_i}\chainedlink{\overline{s1}_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.\overline{S1} 
\  + \ \startchain{\overline{s1}_i}\chainedlink{\overline{s1}_o}{\noact}\chainedlink{\noact}{\hat{s1}_i}\chainend{\hat{s1}_o}.\overline{S1}\ + \  \link{\hat{s1}_i}{\hat{s1}_o}.\overline{S1}\\
 \\
 \\
 S2 & \defeq &\sum_{h=0}^{2} \sum_{k=0}^{2}  \underbrace{ \startchain{s2_i}\chainedlink{s2_o}{\noact}\chainedlink{\noact}{s2_i}\chainedlink{s2_o}{\noact}\chainend{\noact}...}_{size = h}\underbrace{ \startchain{\tilde{s2}_i}\chainedlink{\tilde{s2}_o}{\noact}\chainedlink{\noact}{\tilde{s2}_i}\chainedlink{\tilde{s2}_o}{\noact}\chainend{\noact}...}_{size = k}\link{\hat{s2}_i}{\hat{s2}_o}.S2\\
&&\ +\\
&&\sum_{h=0}^{2}   \underbrace{ \startchain{s2_i}\chainedlink{s2_o}{\noact}\chainedlink{\noact}{s_i}\chainedlink{s2_o}{\noact}\chainend{\noact}...}_{size = h}\link{\hat{s2}_i}{\hat{s2}_o}.\overline{S2}

% \ + \ \link{\hat{a}_i}{\hat{a}_o}.\overline{P_s})
 \\
 \overline{S2}& \defeq &\sum_{h=0}^{2} \sum_{k=1}^{2} \underbrace{ \startchain{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\noact}\chainedlink{\noact}{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\noact}\chainend{\noact}...}_{size = h}\underbrace{ \startchain{\tilde{s2}_i}\chainedlink{\tilde{s2}_o}{\noact}\chainedlink{\noact}{\tilde{s2}_i}\chainedlink{\tilde{s2}_o}{\noact}\chainend{\noact}...}_{size = k}.\link{\hat{s2}_i}{\hat{s2}_o}.S2\\
 && \ +\\
 && \sum_{h=0}^{2}  \underbrace{ \startchain{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\noact}\chainedlink{\noact}{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\noact}\chainend{\noact}...}_{size = h}.\link{\hat{s2}_i}{\hat{s2}_o}.\overline{S2}\\
% 
% 
%   S2 &\defeq& \startchain{e2_i}\chainedlink{e2_o}{\noact}\chainedlink{\noact}{\tilde{e2}_i}\chainedlink{\tilde{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.S2 \ + \
% \startchain{e2_i}\chainedlink{e2_o}{\noact}\chainedlink{\noact}{e2_i}\chainedlink{e2_o}{\noact}\chainedlink{\noact}{\tilde{e2}_i}\chainedlink{\tilde{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.S2 \ + \ \startchain{\tilde{e2}_i}\chainedlink{\tilde{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.S2\\
% && + \\
% &&  \startchain{e2_i}\chainedlink{e2_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.\overline{S2} 
%\  + \ \startchain{e2_i}\chainedlink{e2_o}{\noact}\chainedlink{\noact}{e2_i}\chainedlink{e2_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.\overline{S2}\ + \  \link{\hat{e2}_i}{\hat{e2}_o}.\overline{S2}\\
%\\
%\overline{S2} &\defeq& \startchain{\overline{e2}_i}\chainedlink{\overline{e2}_o}{\noact}\chainedlink{\noact}{\tilde{e2}_i}\chainedlink{\tilde{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.S2 \ + \
% \startchain{\overline{e2}_i}\chainedlink{\overline{e2}_o}{\noact}\chainedlink{\noact}{\overline{e2}_i}\chainedlink{\overline{e2}_o}{\noact}\chainedlink{\noact}{\tilde{e2}_i}\chainedlink{\tilde{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.S2 \ + \ \startchain{\tilde{e2}_i}\chainedlink{\tilde{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.S2\\
% && + \\
% &&  \startchain{\overline{e2}_i}\chainedlink{\overline{e2}_o}{\noact}\chainedlink{\noact}{\overline{e2}_i}\chainedlink{\overline{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.\overline{S2} 
%\  + \ \startchain{\overline{e2}_i}\chainedlink{\overline{e2}_o}{\noact}\chainedlink{\noact}{\hat{e2}_i}\chainend{\hat{e2}_o}.\overline{S2}\ + \  \link{\hat{e2}_i}{\hat{e2}_o}.\overline{S2}\\
 \\
 \\
   S3 &\defeq& \startchain{s3_i}\chainedlink{s3_o}{\noact}\chainedlink{\noact}{\tilde{s3}_i}\chainedlink{\tilde{s3}_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.S3 \ + \
 \startchain{s3_i}\chainedlink{s3_o}{\noact}\chainedlink{\noact}{s3_i}\chainedlink{s3_o}{\noact}\chainedlink{\noact}{\tilde{s3}_i}\chainedlink{\tilde{s3}_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.S3 \ + \ \startchain{\tilde{s3}_i}\chainedlink{\tilde{s3}_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.S3\\
 && + \\
 &&  \startchain{s3_i}\chainedlink{s3_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.\overline{S3} 
\  + \ \startchain{s3_i}\chainedlink{s3_o}{\noact}\chainedlink{\noact}{s3_i}\chainedlink{s3_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.\overline{S3}\ + \  \link{\hat{s3}_i}{\hat{s3}_o}.\overline{S3}\\
\\
\overline{S3} &\defeq& \startchain{\overline{s3}_i}\chainedlink{\overline{s3}_o}{\noact}\chainedlink{\noact}{\tilde{s3}_i}\chainedlink{\tilde{s3}_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.S3 \ + \
 \startchain{\overline{s3}_i}\chainedlink{\overline{s3}_o}{\noact}\chainedlink{\noact}{\overline{s3}_i}\chainedlink{\overline{s3}_o}{\noact}\chainedlink{\noact}{\tilde{s3}_i}\chainedlink{\tilde{s3}_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.S3 \ + \ \startchain{\tilde{s3}_i}\chainedlink{\tilde{s3}_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.S3\\
 && + \\
 &&  \startchain{\overline{s3}_i}\chainedlink{\overline{s3}_o}{\noact}\chainedlink{\noact}{\overline{s3}_i}\chainedlink{\overline{s3}_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.\overline{S3} 
\  + \ \startchain{\overline{s3}_i}\chainedlink{\overline{s3}_o}{\noact}\chainedlink{\noact}{\hat{s3}_i}\chainend{\hat{s3}_o}.\overline{S3}\ + \  \link{\hat{s3}_i}{\hat{s3}_o}.\overline{S3}\\
 \end{array}
 \]

 The code for the rules is as follow.:
 \begin{itemize}
\item Rule 1
 \[
 \begin{array}{lcl}
 R1 &\defeq & \startchain{\silent}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainedlink{\overline{s3}_i}{\noact}\chainedlink{\noact}{\overline{s3}_o}\chainedlink{\tilde{s1}_i}{\noact}\chainedlink{\noact}{\tilde{s1}_o}\chainend{rule2}.R1
\ +\ 
 \startchain{\silent}\chainedlink{\overline{s1}_i}{\noact}\chainedlink{\noact}{\overline{s1}_o}\chainend{rule2}.R1
 \ +\ 
  \startchain{\silent}\chainedlink{s3_i}{\noact}\chainedlink{\noact}{s3_o}\chainend{rule2}.R1\\
 \end{array}
 \]
 \item Rule 2
  \[
 \begin{array}{lcl}
 R2 &\defeq & \startchain{rule2}\chainedlink{s2_i}{\noact}\chainedlink{\noact}{s2_o}\chainedlink{\overline{s1}_i}{\noact}\chainedlink{\noact}{\overline{s1}_o}\chainedlink{\tilde{s3}_i}{\noact}\chainedlink{\noact}{\tilde{s3}_o}\chainend{rule3}.R2\\
 &&+\\
 &&  \startchain{rule2}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainend{rule3}.R2\\
 &&+\\
 &&  \startchain{rule2}\chainedlink{s1_i}{\noact}\chainedlink{\noact}{s1_o}\chainend{rule3}.R2\\
 \end{array}
 \]
 \item Rule 3
  \[
 \begin{array}{lcl}
 R3 &\defeq & \startchain{rule3}\chainedlink{s3_i}{\noact}\chainedlink{\noact}{s3_o}\chainedlink{\overline{s2}_i}{\noact}\chainedlink{\noact}{\overline{s2}_o}\chainedlink{\tilde{s2}_i}{\noact}\chainedlink{\noact}{\tilde{s2}_o}\chainend{stop}.R3\\
 &&+\\
 &&  \startchain{rule3}\chainedlink{\overline{s3}_i}{\noact}\chainedlink{\noact}{\overline{s3}_o}\chainend{stop}.R3\\
 &&+\\
 &&  \startchain{rule3}\chainedlink{s2_i}{\noact}\chainedlink{\noact}{s2_o}\chainend{stop}.R3\\
 \end{array}
 \]
\end{itemize}
Finally, the process $STOP$ is defined as:
\[
STOP\ \defeq \ \startchain{stop}\chainedlink{\hat{s1}_i}{\noact}\chainedlink{\noact}{\hat{s1_o}} \chainedlink{\hat{s2}_i}{\noact}\chainedlink{\noact}{\hat{s2}_o}\chainedlink{\hat{s3}_i}{\noact}\chainedlink{\noact}{\hat{s3}_o}\chainend{\silent}.STOP
\]

Now, let assume that our initial system is composed by only the element $S1$, as follow:
\[
System \ \defeq \ S1|\overline{S2}|\overline{S3}|R1|R2|R3|STOP
\]
Then, it is possible to apply only the first rule. The {\tt link}-calculus transition follows:
\[
\begin{array}{ll}
System & \rightarrow \\
&\startchain{\silent}\chainedlink{s1_i}{s1_i}\chainedlink{s1_o}{s1_o} \chainedlink{\overline{s3}_i}{\overline{s3}_i}\chainedlink{\overline{s3}_o}{\overline{s3}_o}\chainedlink{\tilde{s1}_i}{\tilde{s1}_i}\chainedlink{\tilde{s1}_o}{\tilde{s1}_o}\chainedlink{rule2}{rule2}\chainedlink{\overline{s2}_i}{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\overline{s2}_o}\chainedlink{s1_i}{s1_i}\chainedlink{s1_o}{s1_o}\chainedlink{rule3}{rule3}\chainedlink{\overline{s3}_i}{\overline{s3}_i}\chainedlink{\overline{s3}_o}{\overline{s3}_o}\chainedlink{\overline{s2}_i}{\overline{s2}_i}\chainedlink{\overline{s2}_o}{\overline{s2}_o}\chainedlink{stop}{stop}\\
&
\chainedlink{\hat{s1}_i}{\hat{s1}_i}\chainedlink{\hat{s1}_o}{\hat{s1}_o}\chainedlink{\hat{s2}_i}{\hat{s2}_i}\chainedlink{\hat{s2}_o}{\hat{s2}_o}\chainedlink{\hat{s2}_o}{\hat{s2}_o}\chainedlink{\hat{s3}_i}{\hat{s3}_i}\chainedlink{\hat{s3}_i}{\hat{s3}_i}\chainedlink{\hat{s3}_o}{\hat{s3}_o}\chainend{\silent} \ System
\end{array}
\]
 \end{example}

% Let ${\cal A} = (S, A)$ be a reaction system, and let $t$ be a transition such that $t= ([|{\cal A}|] \xrightarrow{\link{\silent}{\silent}}P')$. By Definition~\ref{def:trans}, we know that
% each rule $a_j \in A$ is encoded by the process $A_j$;
% we denote as the \emph{required elements} by  $A_j$ in $t$ all the elements appearing in the link chain prefix offered by  $A_j$ in $t$.

 \begin{lemma}
 \label{lem:allrules}
  Let ${\cal A} = (S, A)$ be a reaction system, let $P=[|{\cal A}|]$ its c\CNA~translation, and let $t$
be a transition such that  if $t= (P \xrightarrow{\link{\silent}{\silent}} P')$, then
  \begin{enumerate}
  \item for each rule  $a_j \in A$, the corresponding process $A_j = [|a_j|]$,  is involved in the transition $t$;
  \item each process $P_s = [|s|]$, with $s \in S$, participates at most  with one solid link for each rule $A_j$.%, if $s$ is a \emph{required element} of $A_j$ in $t$.
  \end{enumerate}
  \end{lemma}
  \begin{proof}
  We prove the two items separately:
  \begin{enumerate}
\item by Definition~\ref{def:trans}, all the names $r_j$, where $j$ is the counter of the rules in ${\cal A}$, are restricted. Then, all the processes $A_j$ must participate to each transitions.
 \item the code of each rule leaves one virtual link between two solid links of the types $\dots\chainedlink{s_{1i}}{\noact}\chainedlink{\noact}{s_{1o}}\dots$. Then, it derives that the process $P_{s_1}$, that  codes the behaviour of element $s_1$, can participate in the transition with only one solid link $\link{s_{1i}}{s_{1o}}$. In fact, with a number $v$ of virtuals  $v > 1$, 
 $P_{s_1}$ can not participate with a link chain, as by juxtaposing  $\link{s_{1i}}{s_{1o}}$, not valid
 link chains would be generated.
    \end{enumerate}
  \end{proof}

 \begin{proposition}[completeness]
 Let ${\cal A} = (S, A)$ be a reaction system, and let $P=[|{\cal A}|]$ its c\CNA~translation, then if ${\cal A}$ evolves in ${\cal A'} = (S', A)$ in one step, then $\exists$ $P'$ such that $P \rightarrow P'$ and  $P'=[|{\cal A'}|]$.
 \end{proposition}
 \begin{proof}
 The proof proceeds by cases on the possible evolution of each element  $e \in S$, following the execution of one step in the reaction system ${\cal A} = (S, A)$.\\
 %Let  ${\cal A} = (S, A)$ be a reaction system and let $e$ be an element in the system.
 % Then 
  Let ${\cal A'} = (S', A)$ be the evolution of  ${\cal A}$ after one step. There are four cases to be considered: \ 
 1) \ $e \in S$, $e \in S'$; \ 2)\  $e \in S$, $e \notin S'$; \ 3)\  $e \notin S$, $e \notin S'$; \ 4)\  $e \notin S$, $e \in S'$.\\
 case  $e \in S$ and $e \in S'$\\
It means that have been applied some rules  in $A$ producing the element $e$, and also that some rules in $A$ requiring the element $e$ could have been applied. All the rules requiring the absence of $e$ have not been applied.\\
%Then, let $R_i = [e';e'',e]$ be a rule producing $e$ with $e'' \neq e$, and let $i$ be its associated number. \\
It is important to observe that our c\CNA encoding, for each transition, requires the application of all the rules in the system, as the channels $r_j$, with $j \in [1,m]$ are restricted.
Each rule, define a pattern to be satisfied, i.e. each rule inserts as many virtual links  as  the number of reactants, inhibitors, and products, as required by the corresponding RS rule.
It follows that

The code executed by any rule $R_i$ producing $e$ is $R_i \defeq \startchain{rule_i}\chainedlink{e'_i}{\noact}\chainedlink{\noact}{e'_o}\chainedlink{\overline{e''}_i}{\noact} \chainedlink{\noact}{\overline{e''}_o}\chainedlink{\tilde{e}_i}{\noact}\chainedlink{\noact}{\tilde{e}_o} \chainend{rule_{i+1}}.R_i + \dots$, with $e'' \neq e$.\\
The code executed by any  rules $j$ consuming $e$ without producing it is $R_j \defeq \startchain{rule_j}\chainedlink{e_i}{\noact}\chainedlink{\noact}{e_o}\chainedlink{\overline{e'}_i}{\noact} \chainedlink{\noact}{\overline{e'}_o}\chainedlink{\tilde{e''}_i}{\noact}\chainedlink{\noact}{\tilde{e''}_o} \chainend{rule_{j+1}}.R_j + \dots$, with $e',e''\neq e $.\\
The code executed by any rule $j$ not requiring $e$ is $R_j \defeq \startchain{rule_j}\chainedlink{e_i}{\noact}\chainedlink{\noact}{e_o} \chainend{rule_{j+1}}.R_j + \dots$.\\
The code executed by the element $e$ is $P_e \defeq  \underbrace{ \startchain{e_i}\chainedlink{e_o}{\noact}\chainedlink{\noact}{e_i}\chainedlink{e_o}{\noact}\chainend{\noact}...}_{size = h}\underbrace{ \startchain{\tilde{e}_i}\chainedlink{\tilde{e}_o}{\noact}\chainedlink{\noact}{\tilde{e}_i}\chainedlink{\tilde{e}_o}{\noact}\chainend{\noact}...}_{size = k}\link{\hat{e}_i}{\hat{e}_o}.P_s + \dots$, where $h$ is the number of  rules requiring $e$ and $k$ is the number of active rules producing $e$.
Then we observe that, together with the process $STOP$, the resulting processes are $R_i$, $R_j$, $P_e$ and $STOP$, i.e. the element $e$ is produced, because at least a rule $R_i$ produces $e$. \\
The other cases are similar. 
\end{proof}

 \begin{corollary}
  Let ${\cal A} = (S, A)$ be a reaction system, and let $P=[|{\cal A}|]$ its {\tt link}-calculus translation, then if ${\cal A}$ evolves in ${\cal A'} = (S', A)$  then $\exists$ $P'$ such that $P \rightarrow^* P'$ and  $P'=[|{\cal A'}|]$.  
 \end{corollary}

 \begin{proposition}[correctness]
 Let $P$ a {\tt link}-calculus process such that exists a reaction system ${\cal A}$ such that $P=[|{\cal A}|]$, then if $P \rightarrow P'$, then $\exists$ ${\cal A'}$ such that ${\cal A}$  evolves in ${\cal A'}$ in one step and $P'=[|{\cal A'}|]$.
 \end{proposition}

 \begin{corollary}
  Let $P$ a {\tt link}-calculus process such that exists a reaction system ${\cal A}$ such that $P=[|{\cal A}|]$, then if $P \rightarrow^* P'$, then $\exists$ ${\cal A'}$ such that ${\cal A}$  evolves in ${\cal A'}$ and $P'=[|{\cal A'}|]$.
 \end{corollary}