### 2020-07-27 fragments/old/06-timedComputation.tex

% !TEX root = ./main_RS2L.tex

\section{Timed Computation}

In the classical Reaction Systems, the actual state is the union of the elements resulting from the 
previous reaction, and of the elements that the context can provide.

The context can provides elements at certain points along the computation. Here we want to 
simulate this behaviour by introducing a technique that we call \emph{timed programmed computation}.
We introduce a two step computation: one step correspond to the transition simulating the reaction in a reaction system, the other step simulates the time passing.
Within this framework, we can introduce macros for delaying (and programming in advance) the entry
of those elements that in the Reaction Systems are provided by the context at certain points.
The macros $MACRO_j(P_s,n)$ are univocally numbered, and takes as parameters the program  $P_s$(simulating the $s$ element in the RS), and the delay $n$. Then, after $n$ computation steps  $MACRO_j(P_s,n)$  makes program $\overline{P}_s$ to turn into $P_s$, if is not already in the state $P_s$.\\
The c\CNA code for $MACRO_j(P_s,n)$ follows:
\[
\begin{array}{lcl}
MACRO_j(P_s,n) &\defeq & \link{t_j}{t_{j+1}}.MACRO_j(P_s,n-1)\\
MACRO_j(P_s,0) & \defeq & \link{t_j}{act_{s_i}}.\nil | \link{act_{s_o}}{t_{j+1}}.FWD_j\\
FWD_j & \defeq & \link{t_j}{t_{j+1}}.FWD_j\\
\end{array}
\] 
where $j = w$ and $w$ is the number of macro processes.\\
Then, we add an an additional process $START$ that decides the alternation between a RS computational step and a step simulates the time passing.

\[
\begin{array}{lcl}
START& \defeq &\link{\silent}{r_1}.\link{\silent}{t_1}.START
\end{array}
\]

Finally, we have to slightly  modify the code of $P_s$ processes, in both  active ($P_s$) or inactive ($\overline{P_s}$) state, corresponding to each element $s \in S$ of a RS system:
\[
\begin{array}{lcl@{\hspace {1cm}} lcl}
P_s &\defeq &   \dots  & \overline{P}_s &\defeq &   \dots \\
 && +  & && +\\
 && \link{act_{s_i}}{act_{s_o}}.P_s &
  && \link{act_{s_i}}{act_{s_o}}.P_s
\end{array}
\]