### 2020-08-27 TCS_v3/AnswerToReviewer.txt
ringraziamenti progetti e indice j nella prova

Dear Editor and Referees,

we thank you very much for your detailed comments and remarks, which greatly helped us to correct and improve our presentation. 
We updated the manuscript following the advices of the referees. Please find below a detailed description of the way in which 
the items pointed out by the referees were addressed in the revised version.

Best regards,
Linda Brodo, Roberto Bruni, Moreno Falaschi

-------------------------

Manuscript Number: TCS-D-20-00267R1
Manuscript Title: A process algebraic approach to reaction systems.
Author(s): Linda Brodo, Researcher; Roberto Bruni, Professor; Moreno Falaschi, Professor
Submitted to: Theoretical Computer Science

-------------------------
Reviewer #1: I think my original comments have been addressed satisfactorily by the authors, but I still have some remarks:

p.6, l.12 - replace "set of entities" by "sets of entities"
> OK. Done

p.7, l.42 - replace "where v_i is a link chain" by "where v is a link chain" 
as in the syntax from the previous line you use v.P and not v_i.P
> OK. Done

p.8, l.13, (Res) rule - make sure all a's have same font as it seems the one in (\nu a)P' is different
> OK. Done

p.8, l.30 - replace "Fig.1" by "Fig. 1"
> OK. Done

p.12, l.35 (item 2 of Lemma 12) - replace "for each reaction" by "each reaction"
> OK. We have corrected the formulation.

p.12, l.51 - in Def. 13 shouldn't be P \xrightarrow{(\nu names)v} P' instead of P \xrightarrow{v}P' ?  Only starting on p.15, l.52 you say "In this section, for simplicity, we shall often omit topmost restrictions (\nu names)". Thus, either move this assumption earlier or till Section 5 keep using (\nu names) on \xrightarrow
>  No, Def.13 is correct: you can imagine that P contains no restrictions and therefore P \xrightarrow{v}P' is such that all the names in v are visible. Then (\nu names)P \xrightarrow{(\nu names)v} (\nu names)P' but in the label (\nu names)v all the names are transformed to silent actions \tau.

p.13, l.56 - (\nu names) missing from \xrightarrow label
>  No, see above.

p.17, l.49 - replace "(starting at cxt and ending at p1" by "(starting at cxt and ending at p1)", namely add missing bracket
> OK. Done

p.33, l.41 - shouldn't be "j \geq 1" (as assumed on p.33, l.38) instead of "j \geq 0"?
> OK. Done

p.33, l.47 and l.49 - in the proof of the base case "j=1", instead of "j-1" shouldn't be "j"? This is due to the fact that Prop. 16 you mention you have the case j=1 and also on p.33 on l.52 you got \gamma^j and not \gamma^{j-1}.
>  We imagine you meant "in the proof of the inductive case". Yuo are right, we have slightly expanded the proof to make it clear the application of the inductive hypothesis.